---
title: "p8105_hw1_sz3319"
author: "Shiyu Zhang"
date: "2024-09-21"
output: html_document

---


## Library Packages for Coding

```{r library package, warning = FALSE, message = FALSE}
# Library necessary packages
library(ggplot2)
library(tidyverse)
```

## Problem 1

### a) Load the Dataset

```{r setup and data load, warning = FALSE, message = FALSE}
# Load "penguins" dataset
library(palmerpenguins)
data("penguins", package = "palmerpenguins")
```

### b) Describe **penguins** Dataset

```{r, message = FALSE, results = 'hide'}
# Look at data
tail(penguins, 5)
skimr::skim(penguins)
```

The `penguins` dataset contains data on 


### c) Creat and Save a Scatterplot

```{r scatterplot, echo = TRUE}
library(ggplot2)
ggplot(data = penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point(na.rm = TRUE) +
  labs(title = "Flipper Length vs Bill Length in Different Species of Penguins",
       x = "Bill Length (mm)",
       y = "Flipper Length (mm)") +
  theme_minimal()
```

```{r save plot}
ggsave("penguins_scatter_plot.pdf", height = 4, width = 6)
```


## Problem 2

### a) Create the Data

```{r data}
# For reproducibility
set.seed(1234)

# a random sample of size 10 from a standard Normal distribution
normal_sample <- rnorm(10)

# a logical vector indicating whether elements of the sample are greater than 0
logical_vector <- normal_sample > 0

# a character vector of length 10
character_vector <- sample(letters, 10, replace = TRUE)

# a factor vector of length 10, with 3 different factor “levels”
factor_vector <- factor(sample(c("A", "B", "C"), 10, replace = TRUE))
```


### b) Create a Data Frame

```{r df}
# create data frame "df"
df <- tibble(
  normal = normal_sample,
  logical = logical_vector,
  character = character_vector,
  factor = factor_vector
)
```

### c) Calculate the mean of different variables

```{r calculate_means, warning = FALSE, message = FALSE}
# calculate the mean of normal sample
mean_normal <- mean(pull(df, normal))
mean_normal

#calculate the mean of 
mean_logical <- mean(pull(df, logical))
mean_logical

#calculate the mean of 
mean_character <- mean(pull(df, character))
mean_character

#calculate the mean of 
mean_factor <- mean(pull(df, factor))
mean_factor
```

The result shows that:

* The mean of the **normal sample** is `r mean(pull(df, normal))`.
* The mean of **logical vectors** is `r mean(pull(df, logical))`, for the `TRUE` is `1` and `FALSE` is `0`.
* The mean of characters and factors can't be calculated for they are not numeric data.




